---
title: "PH1 HES Training - Inclusion 1 and filtering DF"
author: "S Jarvis"
date: "21/01/2022"
output: html_document
---

```{r HiQ Logo,  echo=FALSE, message=FALSE, warning=FALSE}
### DO NOT REMOVE ###
htmltools::img(src = "https://www.healthiq.co.uk/images/healthiq-logo.svg", 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px;')
knitr::opts_chunk$set(echo = TRUE)
```


```{r Load libraries}
### Load basic database libraries
library(RPostgreSQL)
library(DBI)
# Add other packages here:
library(scales)
library(tidyverse)
library(lubridate) #package for date
library(stringr)
library(magrittr)
library(openxlsx)
library(dbplyr)
library(RPostgres)
library(odbc)
library(janitor)
library(kableExtra)
library(ggplot2)
library(AnalystsHIQ)
library(data.table)
library(DT)

```

```{r Format Table Function}
Formatted_Table <- function(df){
  
  df %>%  
    datatable(extensions = 'Buttons',
                    rownames = FALSE,
                    filter = "top",
                 options = list(dom = 'Blfrtip',
                                buttons = c('copy', 'csv', 'excel'),
                                lengthMenu = list(c(10,25,50,-1),
                                                  c(10,25,50,"All"))))
}
```


```{r Connection set-up, echo=FALSE, include=FALSE}
## This chunk should NEVER be included in anything that goes outside of Health iQ.
# This function sets up your connection to the DB, please edit as appropriate
hes_con <- DBI::dbConnect(
  odbc::odbc(), 
  Driver = "Amazon Redshift (x64)",
  ### CHANGE WHICH DATABASE YOU WANT
  Database = "hesdata_processed_hiq",
  Server = "172.31.32.171",
  Port = 5439,
  ### CHANGE THIS TO YOUR DB USERNAME
  UID = Sys.getenv("db_user"),
  ### HiQ DB password (do not put your password in plain text here!)
  PWD = keyring::key_get("cluster_access"))
```

### General Codes

This uploads the general codes. 
```{r Load Codes}
schema <- "ph1_hes_training"

IP_Full <- tbl(hes_con, in_schema("public","inpatient_mapping"))
OP_Full <- tbl(hes_con, in_schema("public","outpatient_mapping"))

codes <- tbl(hes_con, in_schema(schema, "lookup"))

Start_Date <- ymd("2016-04-01")
End_Date <- ymd("2020-12-31")

inclusion_1 <- filter(codes, category == "Inclusion Part 1") %>% pull(code)
inclusion_2 <- filter(codes, category == "Inclusion Part 2") %>% pull(code)
IBD <- filter(codes, category == "IBD") %>% pull(code)

codes %>% collect() %>% Formatted_Table()
```

## Cohort

Inclusion criteria
•	Universal PH1 cohort:
o	ICD-10 code for PH1 (E74.8) in any position within Inpatient or Outpatient within the study period.
o	Any coding for kidney stones (N20.0) or nephrocalcinosis (E83.5) in Inpatient or Outpatient at any point in the available data.
•	Adult PH1 cohort:
o	Patients within the Universal PH1 cohort aged 18 or older at the last date within the study period.
•	Non-adult PH1 cohort
o	Patients within the Universal PH1 cohort aged <18 years at the last date within the study period.

3.3.3.2	Exclusion criteria
The exclusion criteria for all cohorts are the following, as they constitute basic data quality issues:
•	Age missing or recorded as >120
•	Completely missing sex information, not including the “Not known” and “Not Specified” categories.


### Patients Diagnosed With Ph1


```{r Ph1 Diagnosis}
# Firstly we want to identify patients E748 within our study period. E748 is "Other Specified Disorders of Carbohydrate Metabolism"

#Inpatient
Inc1_Pats_IP <- IP_Full %>%
  filter(between(disdate, Start_Date, End_Date)) %>%
  filter_diags(inclusion_1, collapse = TRUE) %>%
  group_by(hesid) %>%
  mutate(indexdate = min(disdate), 
         age = min(if_else(endage >= 7000, 0, endage))) %>%
  ungroup %>%
  distinct(hesid, indexdate, age) %>%
  compute()

count(Inc1_Pats_IP)

#Outpatient
Inc1_Pats_OP <- OP_Full %>%
  filter(between(apptdate, Start_Date, End_Date)) %>%
  filter_diags(inclusion_1, collapse = TRUE) %>%
  group_by(hesid) %>%
  mutate(
    indexdate = min(apptdate)) %>%
  ungroup %>%
  distinct(hesid, indexdate) %>%
  compute()

count(Inc1_Pats_OP)


Inc1_Pats <- union(Inc1_Pats_IP,Inc1_Pats_OP) %>%
  group_by(hesid) %>%
  mutate(indexdate = min(indexdate),
         age = min(age)) %>%
  ungroup %>%
  distinct(hesid, indexdate, age) %>%
  compute_d(schema, "inc1_patients")

count(Inc1_Pats)

```

### Create Filtered dataframes

This filters the whole datasets to only include Ph1 Patients.

```{r Filter Datasets}
#Make a cut of data with just patients with ph1 for speed
IP_Filtered <- IP_Full %>%
  semi_join(Inc1_Pats, by = "hesid") %>%
  compute_d(schema, "inpatient_mapping_ph1_pat")
OP_Filtered <- OP_Full %>%
  semi_join(Inc1_Pats, by = "hesid") %>%
  compute_d(schema, "outpatient_mapping_ph1_pat")
AE_Filtered <- tbl(hes_con, in_schema("public","aande_mapping")) %>%
  semi_join(Inc1_Pats, by = "hesid") %>%
  compute_d(schema, "aande_mapping_ph1_pat")

IP_Pivoted <- IP_Filtered %>%
  pivot_longer(starts_with(c("diag", "opertn")), names_to = "code_type", values_to = "code") %>%
  filter(code != "") %>%
  mutate(prim = case_when(code_type %in% c("diag_01", opertn_01) ~ 1,
                               TRUE ~ 0), 
         code_type = case_when(str_detect(code_type, "diag") ~ "diag",
                               str_detect(code_type, "oper") ~ "oper",
                                          TRUE ~ "none")) %>%
  compute_d(schema, "inpatient_mapping_ph1_pat_pivot")

OP_Pivoted <- OP_Filtered %>%
  pivot_longer(starts_with(c("diag", "opertn")), names_to = "code_type", values_to = "code") %>%
  filter(code != "") %>%
  mutate(prim = case_when(code_type %in% c("diag_01", opertn_01) ~ 1,
                               TRUE ~ 0), 
         code_type = case_when(str_detect(code_type, "diag") ~ "diag",
                               str_detect(code_type, "oper") ~ "oper",
                                          TRUE ~ "none")) %>%
  compute_d(schema, "outpatient_mapping_ph1_pat_pivot")
```