---
title: "PH1 HES Training - Cohort Creation"
author: "S Jarvis"
date: "12/11/2021"
output: html_document
---

```{r HiQ Logo,  echo=FALSE, message=FALSE, warning=FALSE}
### DO NOT REMOVE ###
htmltools::img(src = "https://www.healthiq.co.uk/images/healthiq-logo.svg", 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px;')
knitr::opts_chunk$set(echo = TRUE)
```


```{r Load libraries}
### Load basic database libraries
library(RPostgreSQL)
library(DBI)
# Add other packages here:
library(scales)
library(tidyverse)
library(lubridate) #package for date
library(stringr)
library(magrittr)
library(openxlsx)
library(dbplyr)
library(RPostgres)
library(odbc)
library(janitor)
library(kableExtra)
library(ggplot2)
library(AnalystsHIQ)
library(data.table)
library(DT)

```

```{r Format Table Function}
Formatted_Table <- function(df){
  
  df %>%  
    datatable(extensions = 'Buttons',
                    rownames = FALSE,
                    filter = "top",
                 options = list(dom = 'Blfrtip',
                                buttons = c('copy', 'csv', 'excel'),
                                lengthMenu = list(c(10,25,50,-1),
                                                  c(10,25,50,"All"))))
}
```


```{r Connection set-up, echo=FALSE, include=FALSE}
## This chunk should NEVER be included in anything that goes outside of Health iQ.
# This function sets up your connection to the DB, please edit as appropriate
hes_con <- DBI::dbConnect(
  odbc::odbc(), 
  Driver = "Amazon Redshift (x64)",
  ### CHANGE WHICH DATABASE YOU WANT
  Database = "hesdata_processed_hiq",
  Server = "172.31.32.171",
  Port = 5439,
  ### CHANGE THIS TO YOUR DB USERNAME
  UID = Sys.getenv("db_user"),
  ### HiQ DB password (do not put your password in plain text here!)
  PWD = keyring::key_get("cluster_access"))
```

## Project Description

Primary Hyperoxaluria (PH) constitutes a group of rare inherited disorders of the liver characterised by the overproduction of oxalate, an end-product of metabolism. High levels of oxalate are toxic because oxalate cannot be broken down by the human body and accumulates in the kidneys. Primary Hyperoxaluria Type 1 (PH1) is the most common and severe form, accounting for 70-90% of all PH cases. The client have a subcutaneously administered drug, investigational RNAi therapeutic targeting glycolate oxidase in development for the treatment of PH1. 

The client would like to understand the English PH1 patient population, specifically their demographics, their regional spread and understand their common comorbidities, treatment specialities, and procedures. Additionally, the client would also like to understand the renal health of the PH1 population by tracking the incidence of diagnosis codes attributed to renal health as well as the healthcare resource use and associated costs for PH1 patients with CKD.


## Codes

### General Codes
```{r Table setup}
schema <- "ph1_hes_training"

#IP_Full <- tbl(hes_con, in_schema("public","inpatient_mapping"))
#OP_Full <- tbl(hes_con, in_schema("public","outpatient_mapping"))

IP_Filtered <- tbl(hes_con, in_schema(schema, "inpatient_mapping_ph1_pat"))
OP_Filtered <- tbl(hes_con, in_schema(schema, "outpatient_mapping_ph1_pat"))

codes <- tbl(hes_con, in_schema(schema, "lookup"))

Start_Date <- ymd("2016-04-01")
End_Date <- ymd("2020-12-31")

inclusion_1 <- filter(codes, category == "Inclusion Part 1") %>% pull(code)
inclusion_2 <- filter(codes, category == "Inclusion Part 2") %>% pull(code)
IBD <- filter(codes, category == "IBD") %>% pull(code)

#codes %>% collect() %>% Formatted_Table()

```

## Cohort
Cohort definition
Patients shall be included into the Universal PH1 study cohort based upon the presence of ICD-10 diagnosis codes within their available patient record for PH1 (E74.8), and a record of kidney stones (N20.0) and/or nephrocalcinosis (E83.5) at any point within the data. 
Patients from the Universal PH1 cohort will be stratified into the adult and non-adult cohorts based upon their age at the last date of the study period (31/12/2020). The index date of patients will be defined as the date of their first PH1 coding in the data.
The index date for all cohorts shall be the first appearance of the patient in the data set that fulfils the inclusion and exclusion criteria.  The follow up period for all patients shall start from their index date until the last interaction in the dataset, or until the last date in the study period, whichever comes earlier.

### Inclusion criteria
•	Universal PH1 cohort:
o	ICD-10 code for PH1 (E74.8) in any position within Inpatient or Outpatient within the study period.
o	Any coding for kidney stones (N20.0) or nephrocalcinosis (E83.5) in Inpatient or Outpatient at any point in the available data.
•	Adult PH1 cohort:
o	Patients within the Universal PH1 cohort aged 18 or older at the last date within the study period.
•	Non-adult PH1 cohort
o	Patients within the Universal PH1 cohort aged <18 years at the last date within the study period.

### Exclusion criteria
The exclusion criteria for all cohorts are the following, as they constitute basic data quality issues:
•	Age missing or recorded as >120
•	Completely missing sex information, not including the “Not known” and “Not Specified” categories.


## Patients Diagnosed With Ph1


```{r}
# Firstly we want to identify patients E748 within our study period. E748 is "Other Specified Disorders of Carbohydrate Metabolism"

#Inpatient
#Take filtered dataset
Inc1_Pats_IP <- IP_Filtered %>%
  #Filter by dates
  filter(between(disdate, Start_Date, End_Date)) %>%
  #Filter Diagnosis
  filter_diags(inclusion_1, collapse = TRUE) %>%
  mutate(endage = case_when(endage >= 7000 ~ 0,
                             TRUE ~ endage)) %>%
  #Find earliest discharge date and age at index date
  group_by(hesid) %>%
  mutate(indexdate = min(disdate), 
         age = min(if_else(endage >= 7000, 0, endage))) %>%
  ungroup %>%
  distinct(hesid, indexdate, age) %>%
  compute()

#Outpatient
Inc1_Pats_OP <- OP_Filtered %>%
  filter(between(apptdate, Start_Date, End_Date)) %>%
  filter_diags(inclusion_1, collapse = TRUE) %>%
  mutate(apptage = case_when(apptage >= 7000 ~ 0,
                             TRUE ~ as.numeric(apptage))) %>%
  group_by(hesid) %>%
  mutate(indexdate = min(apptdate),
         age = min(apptage)) %>%
  ungroup %>%
  distinct(hesid, indexdate) %>%
  compute()

#Find Patient Index and upload table of patients with inclusion criteria 1
#Join inpatient and outpatient data
Inc1_Pats <- union(Inc1_Pats_IP,Inc1_Pats_OP) %>% collect()
  #Find earliest of the index dates and ages
Inc1_Pats <- Inc1_Pats %>% 
  group_by(hesid) %>%
  mutate(indexdate = min(indexdate,na.rm = TRUE),
         age = min(age, na.rm = TRUE)) %>%
  distinct(hesid, indexdate, age) %>%
  compute_d(schema, "inc1_patients")

```

### Patients Diagnosed With Kidney Stones (N200) or Nephrocalcinosis (E835) In The Study Period

```{r Inclusion Criteria 2}
# Here we're identifying Patients with Kidney Stones or Nephrocalcinosis In IP or OP in the study period
# We'll then join this with our list of E748 Patients to get the universal cohort

#Inpatient
Inc2_Pats_IP <- IP_Filtered %>%
  filter(between(disdate, Start_Date, End_Date)) %>%
  filter_diags(inclusion_2, collapse = TRUE) %>%
  distinct(hesid) %>%
  compute()


#Outpatient
Inc2_Pats_OP <- OP_Filtered %>%
  filter(between(apptdate, Start_Date, End_Date)) %>%
  filter_diags(inclusion_2, collapse = TRUE) %>%
  distinct(hesid) %>%
  compute()

#Create table patients with inclusion 2
Inc2_Pats <- union(Inc2_Pats_IP,Inc2_Pats_OP) %>% 
  compute_d(schema,"inc2_patients")

```


### Remove patients with exclusion criteria

```{r Exclusion and universal cohort}
#Load Inclusion tables
Inc1_Pats <- tbl(hes_con, in_schema(schema, "inc1_patients"))
Inc2_Pats <- tbl(hes_con, in_schema(schema, "inc2_patients"))

# Here we're Merging the 2 Inclusion Criteria to create the Universal Cohort
Inc_Pats <- Inc1_Pats %>%
  inner_join(Inc2_Pats, by = "hesid") %>%
  compute()

#Additional Analysis: Excluding IBD and Crohn's Patients From all cohorts
#Find exclusion criteria in inpatient
Exclusions_IP <- IP_Filtered %>%
  semi_join(Inc_Pats, by = "hesid") %>%
  filter(between(disdate, Start_Date, End_Date)) %>%
  filter_diags(IBD, collapse = TRUE) %>%
  distinct(hesid) %>%
  compute()

#Find exclusion criteria in outpatient
Exclusions_OP <- OP_Filtered %>%
  semi_join(Inc_Pats, by = "hesid") %>%
  filter(between(apptdate, Start_Date, End_Date)) %>%
  filter_diags(IBD, collapse = TRUE) %>%
  distinct(hesid) %>%
  compute()

#Create a table of exclusion criteria
Exclusions <- Exclusions_IP %>% union(Exclusions_OP) %>%
  distinct(hesid)

#Join inclusion 1 and 2 patients together, and drop those with exclusions
Universal_Cohort <- Inc1_Pats %>%
  inner_join(Inc2_Pats, by = "hesid") %>%
  anti_join(Exclusions, by = "hesid") %>%
  compute()

```

### Drop Dead patients

```{r Drop dead patients}
#Find Patients who had a dead 
Dead_Patients <- Universal_Cohort %>%
  inner_join(IP_Filtered, by = "hesid") %>%
  filter(dismeth == 4) %>%
  distinct(hesid) %>%
  compute_d(schema, "dead_patients")

Invalid_Age <- Universal_Cohort %>%
  anti_join(Dead_Patients, by = "hesid") %>%
  filter(is.na(age)) %>%
  compute()

#Drop patients who died.
Universal_Patients <- Universal_Cohort %>%
  anti_join(Dead_Patients, by = "hesid") %>%
  anti_join(Invalid_Age, by = "hesid") %>%
  rename(index_age = age) %>%
  compute_d(schema,"universal_cohort")

```



### Assign cohorts

```{r}

#Adult PH1 Cohort
Adult_Cohort_Final <- Universal_Patients %>%
  filter(index_age >= 18)%>%
  compute

#Non-Adult PH1 Cohort
Non_Adult_Cohort_Final <- Universal_Patients %>%
  filter(index_age < 18)%>%
  compute

#Creating an overall cohort object
Overall_Cohort <- union_all(
  #Mark the whole universal cohort as being in a single cohort
  Universal_Patients %>%
    mutate(cohort = 0),
  #Add adult cohort
  Adult_Cohort_Final %>%
    mutate(cohort = 1)
  )%>%
    union_all(
  #Add Non-adult cohort
  Non_Adult_Cohort_Final %>%
    mutate(cohort = 2)
  ) %>%
  compute_d(schema,"overall_cohort")

```

#### Checking The Number Of Patients In Each Cohort

```{r}
#Count members of the each cohort
Cohort_Count <- tbl(hes_con, in_schema(schema, "overall_cohort")) %>%
  distinct(hesid, cohort) %>%
  count(cohort) %>%
  arrange(cohort) %>%
  rename(patients = n, condition = cohort) %>%
  mutate(condition = case_when(condition == 0 ~ "Universal Cohort (did not die and valid age)",
                               condition == 1 ~ "Adult at end of study period",
                               TRUE ~ "Non-adult at end of study period")) %>%
  collect %>%
  mutate(condition = factor(condition, c("Adult at end of study period", "Non-adult at end of study period", "Universal Cohort (did not die and valid age)")))

```

## Display: Cohort Counts
```{r}
#Create patient funnel
data.frame(condition = c("PH1", "PH1 & Kidney Stones/Nephrocalcinosis", "With IBD/Chron's", "Without IBD/Chron's", "Died", "Invalid Age", "Alive at End"), 
           patients = c(Inc1_Pats %>% count() %>% pull(n), 
                                  Inc2_Pats %>% count() %>% pull(n), 
                                  Exclusions  %>% count() %>% pull(n), 
                                  Universal_Cohort %>% count() %>% pull(n),
                                  Dead_Patients %>% count() %>% pull(n),
                                  Invalid_Age %>% count() %>% pull(n),
                                  Universal_Patients %>% count() %>% pull(n))) %>% 
  union_all(Cohort_Count) %>%
  mutate(stage = row_number()) %>%
  select(stage, everything()) %>%
  Formatted_Table()
```

## Session Info

In this bit it's always wise to include a session information chunk, that captures specific package and R versions you may be using. This is for reproducibility if a client needs further work much later.

```{r Session Info}
sessioninfo::session_info()
```