---
title: "PH1 HES Training - Descriptive Statistics"
author: "S Jarvis"
date: "12/11/2021"
output: html_document
---

We have agreed to produce the following descriptive statistics
*	Total patients – the number of patients each cohort
*	Total patients with at least one diagnosis of kidney stones – the number of patients within the PH1 cohorts with at least one diagnosis for kidney stones (N20.0).
*	Total patients with at least one diagnosis of nephrocalcinosis – the number of patients within the PH1 cohorts with at least one diagnosis for nephrocalcinosis (E83.5).
*	Age on inclusion – the mean with the standard deviation, median with the IQR, minimum, and maximum age of patients at the last date in the study period.
*	Age distribution on inclusion by decade.
*	Sex – the number and percentage of male, female, and unknown patients
*	Total time in cohort – the total time contributed to the study, calculated from index date to the last available date for each patient, reported in patient-days.
*	Follow-up – the mean, standard deviation, median, IQR, minimum, and maximum number of days of follow-up for each patient.
*	Distribution of Charlson Co-morbidity Index.
*	Ethnicity breakdown
*	Renal health of patients in followup (time post index date)


```{r Load libraries}
### Load basic database libraries
library(RPostgreSQL)
library(DBI)
# Add other packages here:
library(scales)
library(tidyverse)
library(lubridate) #package for date
library(stringr)
library(magrittr)
library(openxlsx)
library(dbplyr)
library(RPostgres)
library(odbc)
library(janitor)
library(kableExtra)
library(ggplot2)
library(AnalystsHIQ)

```

```{r Connection set-up, echo=FALSE, include=FALSE}
## This chunk should NEVER be included in anything that goes outside of Health iQ.
# This function sets up your connection to the DB, please edit as appropriate
hes_con <- DBI::dbConnect(
  odbc::odbc(), 
  Driver = "Amazon Redshift (x64)",
  ### CHANGE WHICH DATABASE YOU WANT
  Database = "hesdata_processed_hiq",
  Server = "172.31.32.171",
  Port = 5439,
  ### CHANGE THIS TO YOUR DB USERNAME
  UID = "seth", 
  ### HiQ DB password (do not put your password in plain text here!)
  PWD = keyring::key_get("database_access"))
```



## Descriptive Statistics
3.3.4.1	Descriptive Statistics:
•	Total patients – the number of patients with an index date during the study period
•	Total patients with at least one diagnosis of kidney stones – the number of patients within the PH1 cohorts with at least one diagnosis for kidney stones (N20.0).
•	Total patients with at least one diagnosis of nephrocalcinosis – the number of patients within the PH1 cohorts with at least one diagnosis for nephrocalcinosis (E83.5).
•	Age on inclusion – the mean, standard deviation, median, IQR, minimum, and maximum age of patients at the last date in the study period.
•	Age distribution on inclusion by decade.
•	Sex – the number and percentage of male, female, and unknown patients
•	Total time in cohort – the total time contributed to the study, calculated from index date to the last available date for each patient, reported in patient-days.
•	Follow-up – the mean, standard deviation, median, IQR, minimum, and maximum number of days of follow-up for each patient.
•	Distribution of Charlson Co-morbidity Index.
•	Ethnicity breakdown, as defined by NHS Digital:
o	White 
o	Black - Caribbean 
o	Black - African 
o	Black - Other 
o	Indian 
o	Pakistani 
o	Bangladeshi 
o	Chinese 
o	Any other ethnic group 
o	Not given 
o	Not known



### General Codes
```{r}

schema <- "ph1_hes_training"

Start_Date <- ymd("2015-04-01")
End_Date <- ymd("2020-12-31")

IP_Filtered <- tbl(hes_con, in_schema(schema, "inpatient_mapping_ph1_pat")) %>%
  filter(between(disdate, Start_Date, End_Date))
OP_Filtered <- tbl(hes_con, in_schema(schema, "outpatient_mapping_ph1_pat")) %>%
  filter(between(apptdate, Start_Date, End_Date))


IP_Pivoted <- tbl(hes_con, in_schema(schema, "inpatient_mapping_ph1_pat_pivot")) %>%
  filter(between(disdate, Start_Date, End_Date))

OP_Pivoted <- tbl(hes_con, in_schema(schema, "outpatient_mapping_ph1_pat_pivot")) %>%
  filter(between(apptdate, Start_Date, End_Date))


Overall_Cohort <- tbl(hes_con, in_schema(schema,"overall_cohort"))

codes <- tbl(hes_con, in_schema(schema, "lookup"))

cci_codes <- tbl(hes_con, in_schema(schema, "lookup_cci_icd10_ammended"))

Ethnicity_Lookup <- tbl(hes_con, in_schema("ph1_hes_training", "ethnicity_lookup"))

codes %>% kable %>% kable_styling()
```



### Calculate Age
•	Age on inclusion – the mean, standard deviation, median, IQR, minimum, and maximum age of patients at their index date.
•	Age distribution on inclusion by decade.

The Age of patients should be calculated at the last date of the study period. We calculated this in the cohort creation step


```{r Age}
#Categorise age at index
Age_With_Cat <- Overall_Cohort %>%
  mutate(age_cat = case_when(index_age < 1 ~ "<1",
                             between(index_age, 1, 20) ~ as.character(round(index_age, 0)),
                             index_age < 31 ~ "21-30",
                             index_age < 41 ~ "31-40",
                             index_age < 51 ~ "41-50",
                             index_age < 61 ~ "51-60",
                             index_age < 71 ~ "61-70",
                             index_age < 81 ~ "71-80",
                             index_age < 91 ~ "81-90",
                             index_age < 101 ~ "91-100",
                             TRUE ~ ">100")) %>%
  collect()

#Find Summary Stats
Age_Summary <- Age_With_Cat %>%
  #Group age by cohort and calculate summary stats
  group_by(cohort, .drop = FALSE) %>%
    summarise(mean = mean(index_age),
              median = median(index_age),
              stdev = sd(index_age),
              maximum = max(index_age),
              minimum = min(index_age)) %>%
  ungroup() %>%
  column_to_rownames("cohort") %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("stat")

#Turn age into a category
Age_Cat_Counts <- Age_With_Cat %>%
  mutate(age_cat = factor(age_cat, levels = c("<1","1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", 
                                              "21-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90", "91-100", ">100"), ordered = TRUE)) %>%
  group_by(cohort, age_cat, .drop = FALSE) %>%
    summarise(counts = n(),
              .groups = "keep") %>%
  ungroup() %>%
  pivot_wider(names_from = cohort, values_from = counts) %>%
  rename(stat = age_cat)


```


### Gender

•	Sex – the number and percentage of male, and female patients

```{r Gender}
#Find all distinct Gender records in IP
Gender_IP <- Overall_Cohort %>%
  inner_join(IP_Filtered, by = "hesid") %>%
  filter(!is.na(sex)) %>%
  mutate(dataset = 1, sex = as.character(sex)) %>%
  distinct(dataset, hesid, sex) %>%
  compute

#Find all distinct gender records in OP
Gender_OP <- Overall_Cohort %>%
  inner_join(OP_Filtered, by = "hesid") %>%
  filter(!is.na(sex)) %>%
  mutate(dataset = 2) %>%
  distinct(dataset, hesid, sex) %>%
  compute

#Find the minimum dataset they have a record of gender in
Gender <- Gender_IP %>% 
  union(Gender_OP) %>% 
  group_by(hesid) %>%
    filter(dataset == min(dataset)) %>%
  ungroup %>%
  inner_join(Overall_Cohort, by = "hesid") %>%
  compute

#Get the counts of gender by cohort
Gender_Stat <- Gender %>%
  count(cohort, sex) %>%
  mutate(count = as.numeric(n)) %>%
  select(-n) %>%
  collect() %>%
  #Find percentage of total
  group_by(cohort) %>%
    mutate(percentage = count / sum(count)) %>%
  ungroup() %>%
  #Rearrange dataset
  pivot_longer(count:percentage, names_to = "stat", values_to = "value") %>%
  pivot_wider(names_from = cohort, values_from = value) %>%
  arrange(sex) %>%
  rename(category = sex)

```

### Time in cohort
•	Total time in cohort – the total time contributed to the study, calculated from index date to the last available date for each patient, reported in patient-days.

```{r Time in Cohort}
#Find last available 
Cohort_Time <- IP_Filtered %>%
  distinct(hesid, disdate) %>%
  union_all(OP_Filtered %>%
            select(hesid, disdate = apptdate) %>%
            distinct(hesid, disdate)) %>%
  #Find maximum observation
  group_by(hesid) %>%
    summarise(max_obs = max(disdate, na.rm = TRUE)) %>%
  ungroup() %>%
  #Join to cohort
  inner_join(Overall_Cohort, by = "hesid") %>%
  select(hesid, cohort, indexdate, max_obs) %>%
  #Get time in cohort in days and months
  mutate(time_in_cohort = max_obs - indexdate,
         months_in_cohort = time_in_cohort/30) %>%
  compute()

#Calculate summary stats by cohort
Cohort_Time_Stats <- Cohort_Time %>%
  collect() %>%
  group_by(cohort) %>%
    summarise(total_time = sum(time_in_cohort),
              mean = mean(months_in_cohort),
              median = median(months_in_cohort),
              stdev = sd(months_in_cohort),
              maximum = max(months_in_cohort),
              minimum = min(months_in_cohort)) %>%
  ungroup() %>%
  column_to_rownames("cohort") %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("stat")
```

### Ethnicity Breakdown
•	Ethnicity breakdown, as defined by NHS Digital:
o	White 
o	Black - Caribbean 
o	Black - African 
o	Black - Other 
o	Indian 
o	Pakistani 
o	Bangladeshi 
o	Chinese 
o	Any other ethnic group 
o	Not given 
o	Not known


```{r Ethnicity}

#Find ethnicity records in IP
Ethnicity_IP <- Overall_Cohort %>%
  inner_join(IP_Filtered, by = "hesid") %>%
  #find time from index
  mutate(index_time = abs(disdate-indexdate)) %>%
  filter(!is.na(ethnos)) %>%
  #find the event closest to index
  group_by(hesid) %>%
    filter(index_time == min(index_time)) %>%
  ungroup %>%
  mutate(dataset = 1) %>%
  distinct(hesid, ethnos, dataset, index_time) %>%
  compute
  
#Find ethnicity records in OP
Ethnicity_OP <- Overall_Cohort %>%
  inner_join(OP_Filtered, by = "hesid") %>%
  mutate(index_time = abs(apptdate-indexdate)) %>%
  filter(!is.na(ethnos)) %>%
  group_by(hesid) %>%
  filter(index_time == min(index_time)) %>%
  ungroup %>%
  mutate(dataset = 2) %>%
  distinct(hesid, ethnos, dataset, index_time) %>%
  compute

#Union ethnicity results
Ethnicity <- union(Ethnicity_IP,Ethnicity_OP) %>%
  filter(!is.na(ethnos)) %>%
  #Join ethnicity lookups 
  inner_join(Ethnicity_Lookup, by = "ethnos") %>%
  #Tag any which are not known as not known
  mutate(known = case_when(category %in% c("Not known", "Not Given") ~ 0,
                          TRUE ~ 1)) %>%
  #Find the minimum dataset and the maximum known for each thing
  group_by(hesid) %>%
    filter(dataset == min(dataset)) %>%
    filter(known == max(known)) %>%
  ungroup() %>%
  distinct(hesid, category) %>%
  group_by(hesid, category) %>%
    mutate(count = n()) %>%
  ungroup() %>%
  mutate(category = case_when(count >= 2 ~ "Not known",
                              TRUE ~ category)) %>%
  distinct(hesid, category) %>%
  inner_join(Overall_Cohort, by = "hesid") %>%
  compute()

#Get counts of each type of ethnicity
Ethnicity_Stats <- Ethnicity %>%
  collect() %>%
  #Turn ethnicity into a factor to amke it easy to arrange
  mutate(category = factor(category, levels = c("White", "Black - Caribbean", "Black - African", "Black - Other", "Indian", "Pakistani", "Bangladeshi", "Chinese", "Any other ethnic group", "Not Given", "Not known"))) %>%
  #Count ethnicity by cohort and don't drop any missing ones
  group_by(cohort, category, .drop = FALSE) %>%
    summarise(counts = n()) %>%
  ungroup() %>%
  #Pivot ethnicity wider by cohort
  pivot_wider(names_from = cohort, values_from = counts) %>%
  mutate(across(where(is.integer), as.numeric),
         across(where(is.numeric), coalesce, 0)) %>%
  arrange(category) %>%
  rename(stat = category)

```


### Calculate CCI
•	Distribution of Charlson Co-morbidity Index.

```{r CCI}
#Find CCI events in IP dataset
CCI_IP <- IP_Pivoted %>%
  filter(code_type == "diag") %>%
  mutate(code = str_sub(code, 1, 4)) %>%
  inner_join(cci_codes, by = c("code" = "icd10")) %>%
  rename(obsdate = disdate) %>%
  distinct(hesid, code, condition, weights, obsdate) %>% 
  mutate(origin = "IP") %>%
  compute()

#Find CCI events in OP dataset
CCI_OP <- OP_Pivoted %>%
  filter(code_type == "diag") %>%
  mutate(code = str_sub(code, 1, 4)) %>%
  inner_join(cci_codes, by = c("code" = "icd10")) %>%
  rename(obsdate = apptdate) %>%
  distinct(hesid, code, condition, weights, obsdate) %>% 
  mutate(origin = "OP") %>%
  compute()

#Join IP and OP data and find distinct conditions and weights per patient
CCI_Union <- CCI_IP %>%
  union_all(CCI_OP) %>%
  inner_join(Overall_Cohort, by = "hesid") %>%
  filter(obsdate <= indexdate) %>%
  distinct(hesid, cohort, condition, weights) %>%
  compute()

#Find CCI score for each patient
CCI_Summaries <- CCI_Union %>%
  group_by(hesid, cohort) %>%
    summarise(cci = sum(weights)) %>%
  ungroup() %>%
  right_join(Overall_Cohort %>% select(hesid, cohort), by = c("hesid", "cohort")) %>%
  mutate(cci = coalesce(as.numeric(cci), 0)) %>%
  mutate(cci_grouping = case_when(cci > 12 ~ ">12",
                                  TRUE ~ as.character(cci))) %>%
  compute()

#Categorise CCI and count patients per category
CCI_Stats <- CCI_Summaries %>%
  collect() %>%
  mutate(cci_grouping = factor(cci_grouping, c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", ">12"), ordered = TRUE)) %>%
  distinct(hesid, cohort, cci_grouping) %>%
  group_by(cohort, cci_grouping, .drop=FALSE) %>%
    summarise(n = as.numeric(n())) %>%
  ungroup() %>%
  pivot_wider(names_from = cohort, values_from = n) %>%
  rename(stat = cci_grouping)
  
```

### Assessing Renal Health Post Index
•	Kidney Stones
•	Dialysis
•	Kidney transplantation
•	Liver transplantation
•	Concomitant liver and kidney transplantation
•	Haematuria
•	Renal failure
•	Urinary tract infection
•	Chronic kidney disease



```{r Comorbidities}

#Find comorbidities in IP
Comorbid_IP <- IP_Pivoted %>%
  filter(code_type == "diag") %>%
  mutate(code = str_sub(code, 1, 4)) %>%
  inner_join((codes %>% filter(code.type == "ICD")), by = "code") %>%
  union_all(IP_Pivoted %>%
    filter(code_type == "oper") %>%
    mutate(code = str_sub(code, 1, 4)) %>%
    inner_join((codes %>% filter(code.type == "OPCS")), by = "code")) %>%
  rename(obsdate = disdate) %>%
  distinct(hesid, code, category, obsdate) %>% 
  mutate(origin = "IP") %>%
  filter(!category %in% c("Inclusion Part 1", "Inclusion Part 2")) %>%
  compute()

#Find comorbidities in OP
Comorbid_OP <- OP_Pivoted %>%
  filter(code_type == "diag") %>%
  mutate(code = str_sub(code, 1, 4)) %>%
  inner_join((codes %>% filter(code.type == "ICD")), by = "code") %>%
  union_all(OP_Pivoted %>%
    filter(code_type == "oper") %>%
    mutate(code = str_sub(code, 1, 4)) %>%
    inner_join((codes %>% filter(code.type == "OPCS")), by = "code")) %>%
  rename(obsdate = apptdate) %>%
  distinct(hesid, code, category, obsdate) %>% 
  mutate(origin = "OP") %>%
  filter(!category %in% c("Inclusion Part 1", "Inclusion Part 2")) %>%
  compute()

#Union IP and OP events
Union_Comorbid <- Comorbid_IP %>%
  union_all(Comorbid_OP) %>%
  distinct(hesid, code, category, obsdate, origin) %>%
  compute()

#This looks for kidney and liver transplants on the same day
Transplant_Concurrent <- Union_Comorbid %>%
  select(-origin, -code) %>%
  filter(category %in% c("Kidney transplantation", "Liver transplantation")) %>%
  pivot_wider(names_from = category, values_from = obsdate) %>%
  rename(kidney = "Kidney transplantation", liver = "Liver transplantation") %>%
  mutate(concurrent = case_when(!is.na(kidney) & kidney == liver ~ kidney,
                               TRUE ~ NA)) %>%
  select(hesid, obsdate = concurrent) %>%
  filter(!is.na(obsdate)) %>%
  mutate(category = "Concurrent transplant") %>%
  compute()

#Find observations after index
Union_Conditions <- Union_Comorbid %>%
  distinct(hesid, obsdate, category) %>% 
  union_all(Transplant_Concurrent) %>%
  right_join(Overall_Cohort %>% distinct(hesid, indexdate), by = "hesid") %>%
  filter(obsdate >= indexdate) %>%
  right_join(Overall_Cohort, by = c("hesid", "indexdate")) %>%
  collect() %>%
  #Set Conditions to factors
  mutate(category = factor(category, levels = c("Kidney Stones", "Dialysis", "Kidney transplantation", "Liver transplantation", "Concurrent transplant", "Haematuria", 
                                                "Renal failure", "Urinary tract infection", "Chronic kidney disease", "CKD Stage 1", "CKD Stage 2", "CKD Stage 3",
                                                "CKD Stage 4", "CKD Stage 5", "Nephrocalcinosis"), ordered = TRUE)) %>%
  group_by(hesid, cohort, indexdate, category, .drop = FALSE) %>%
    summarise(events = n(),
              .groups = "keep") %>%
  ungroup() %>%
  compute()

#Find number of years per cohort to find events per year.
Years_Per_Cohort <- Cohort_Time_Stats %>% 
  filter(stat == "total_time") %>% 
  pivot_longer('0':'2', names_to = "cohort") %>% 
  mutate(value = value/365,
         cohort = as.numeric(cohort)) %>%
  select(-stat)

#Get summary stats
Union_Conditions_Stat <- Union_Conditions %>%
  filter(events > 0) %>%
  inner_join(Years_Per_Cohort, by = "cohort") %>%
  group_by(cohort, category, .drop = FALSE) %>%
    summarise(Total_Events = sum(events),
              Patients = n(),
              Incidence_Per_Year = Total_Events / max(value)) %>%
  ungroup() %>%
  filter(category != "Nephrocalcinosis") %>%
  collect() %>%
  pivot_longer(Total_Events:Incidence_Per_Year, names_to = "stat", values_to = "n") %>%
  pivot_wider(names_from = cohort, values_from = n)
  

```

### Find total patients, and the number with kidney stones and nephrocalcionis

•	Total patients – the number of patients with an index date during the study period
•	Total patients with at least one diagnosis of kidney stones – the number of patients within the PH1 cohorts with at least one diagnosis for kidney stones (N20.0).
•	Total patients with at least one diagnosis of nephrocalcinosis – the number of patients within the PH1 cohorts with at least one diagnosis for nephrocalcinosis (E83.5).



```{r Get cohort inclusion values}
#Count patients per cohort
Overall_Counts <- Overall_Cohort %>% 
  count(cohort) %>%
  mutate(n = as.numeric(n),
         stat = "Total") %>%
  pivot_wider(names_from = cohort, values_from = n) %>%
  collect() %>%
  select(stat, "0", "1", "2")

#Count the number of patients with each of kidney stones and nephrocalcinosis at any point
Inclusion_Counts <- Union_Comorbid %>%
  distinct(hesid, obsdate, category) %>% 
  inner_join(Overall_Cohort, by = "hesid") %>%
  collect() %>%
  filter(category %in% c("Kidney Stones", "Nephrocalcinosis")) %>%
  count(hesid, cohort, category) %>% 
  group_by(cohort, category, .drop = FALSE) %>%
    summarise(n = n()) %>%
  ungroup() %>%
  rename(stat = category) %>%
  pivot_wider(names_from = cohort, values_from = n) %>%
  select(stat, "0", "1", "2")

```

## Join tables into output

```{r Join Tables}
#Join all of the stats tables together
Final_Table <- Overall_Counts %>%
  union_all(Inclusion_Counts) %>%
  union_all(Age_Summary) %>%
  union_all(Age_Cat_Counts) %>%
  union_all(Gender_Stat) %>%
  union_all(Cohort_Time_Stats) %>%
  union_all(Ethnicity_Stats) %>%
  union_all(CCI_Stats) %>%
  union_all(Union_Conditions_Stat)

write.xlsx(Final_Table, "Output/Descriptive_Stats_Table.xlsx")
```

