---
title: "PH1 HES Training - Ranked Analysis"
author: "S Jarvis"
date: "20/01/2022"
output: html_document
---


```{r Load libraries}
### Load basic database libraries
library(RPostgreSQL)
library(DBI)
# Add other packages here:
library(scales)
library(tidyverse)
library(lubridate) #package for date
library(stringr)
library(magrittr)
library(openxlsx)
library(dbplyr)
library(RPostgres)
library(odbc)
library(janitor)
library(kableExtra)
library(ggplot2)
library(AnalystsHIQ)

```

```{r Connection set-up, echo=FALSE, include=FALSE}
## This chunk should NEVER be included in anything that goes outside of Health iQ.
# This function sets up your connection to the DB, please edit as appropriate
hes_con <- DBI::dbConnect(
  odbc::odbc(), 
  Driver = "Amazon Redshift (x64)",
  ### CHANGE WHICH DATABASE YOU WANT
  Database = "hesdata_processed_hiq",
  Server = "172.31.32.171",
  Port = 5439,
  ### CHANGE THIS TO YOUR DB USERNAME
  UID = "seth", 
  ### HiQ DB password (do not put your password in plain text here!)
  PWD = keyring::key_get("database_access"))
```

## Short summary

This is a version of the ranked analysis where the cutoff points are picked manually.

For the Adult and Non-Adult subcohorts we have been asked to produce ranked analysis of:
*	Top 100 Primary Diagnosis (Diagnosis in the primary position)
*	Top 100 Secondary diagnosis (Diagnosis not in the primary position)
*	Top 100 Procedures (In any position)
*	Top 20 treatment specialties of clinicians (found by the tretspef)

### General Codes
```{r}

schema <- "ph1_hes_training"

Start_Date <- ymd("2015-04-01")
End_Date <- ymd("2020-12-31")

IP_Filtered <- tbl(hes_con, in_schema(schema, "inpatient_mapping_ph1_pat")) %>%
  filter(between(disdate, Start_Date, End_Date))
OP_Filtered <- tbl(hes_con, in_schema(schema, "outpatient_mapping_ph1_pat")) %>%
  filter(between(apptdate, Start_Date, End_Date))

IP_Pivoted <- tbl(hes_con, in_schema(schema, "inpatient_mapping_ph1_pat_pivot")) %>%
  filter(between(disdate, Start_Date, End_Date))

OP_Pivoted <- tbl(hes_con, in_schema(schema, "outpatient_mapping_ph1_pat_pivot")) %>%
  filter(between(apptdate, Start_Date, End_Date))

Overall_Cohort <- tbl(hes_con, in_schema(schema,"overall_cohort"))

codes <- tbl(hes_con, in_schema(schema, "lookup"))

Full_Code_Lookup <- tbl(hes_con, in_schema(schema,"full_lookup_names"))

```

```{r Filter Function}

#Create function to filter rows based on the ranks
#Ranking_Dataset is the dataset to be used
#Highest_Row is the highest number of events desired
#Over should be set to true if you are happy with having more than 100 rows, and false if you aren't
#If True it drops the rank which has 
Filter_Rows_Under_Max <- function(Ranking_Dataset, Highest_Row = 100, Over = TRUE){
  
  #Don't filter anything if the highest row is higher than the dataset
  if(Highest_Row > unlist(Ranking_Dataset %>% count() %>% pull(n))){
    print("Your max value is longer than your DF, no rows have been filtered.")
    return(Ranking_Dataset)
  }
  
  #Arrange dataset by appointments then patients and assign row number
  Keep_DF <- Ranking_Dataset %>%
    arrange(ranking) %>%
    mutate(row = row_number()) %>%
    #Group by the rank and find the maximum and minimum row of that rank
    group_by(ranking) %>%
      mutate(max_row = max(row),
             min_row = min(row)) %>%
    ungroup() %>%
    #If Over is true then keep all those which have a minimum row under the ranking level
    #If Over is false keep those with a maximum row under the highest row level.
    mutate(keep = case_when(Over == TRUE & min_row <= Highest_Row ~ TRUE,
                            Over == FALSE & max_row <= Highest_Row ~ TRUE,
                            TRUE ~ FALSE)) %>%
    filter(keep == TRUE) %>%
    select(-row, - min_row, -max_row) %>% 
    compute()
  
  return(Keep_DF)
    
}
```


### Primary Diagnosis
```{r Primary Diag}

# Join cohort to filtered dataset
Primary_Diags <- Overall_Cohort %>%
  inner_join(IP_Pivoted, by = "hesid") %>%
  #Filter to only use primary diagnosis
  filter(prim == 1) %>%
  select(cohort, hesid, disdate, code) %>%
  rename(value = code) %>%
  filter(!is.na(value)) %>%
  #select only the first 4 characters from primary diag column
  mutate(icd10 = substring(value,1,4)) %>%
  #Join to list of all codes for descriptions
  inner_join(Full_Code_Lookup, by = "icd10") %>%
  select(cohort, hesid, icd10, description) %>%
  compute()

#Count number of times each patient has each icd code
Primary_Ranking <- Primary_Diags %>%
  count(hesid, cohort, icd10, description) %>%
  ungroup() %>%
  #Count the number of patients who have each icd code, and the total number of appointments with it
  group_by(cohort, icd10, description) %>%
    summarise(n_pat = as.numeric(n()),
              n_appointments = as.numeric(sum(n))) %>%
  ungroup() %>%
  #Collect before running the section below otherwise it will rank things incorrectly
  collect() %>%
  #Rank ICD codes by number of appointments with them
  group_by(cohort) %>%
    arrange(desc(n_appointments), desc(n_pat)) %>%
    #If prior number of appointments and number of patients is the same set to 0, otherwise set to 1
    #Otherwise set to 1
    mutate(differ_prior = case_when(lag(n_appointments) == n_appointments & lag(n_pat) == n_pat ~ 0,
                                   TRUE ~ 1)) %>%
    #Cumsum the different prior to get rankings
    mutate(ranking = cumsum(differ_prior)) %>%
  ungroup() %>%
  select(-differ_prior)
  

#Find only the adult cohort
Primary_Diag_Adult_Rank <- Primary_Ranking %>%
  filter(cohort == 1) %>%
  #Run function
  Filter_Rows_Under_Max(100, FALSE) %>%
  #Order the rows
  arrange(ranking, desc(n_appointments), desc(n_pat)) %>%
  select(ranking, icd10, description, n_pat, n_appointments)

Primary_Diag_Non_Adult_Rank <- Primary_Ranking %>%
  filter(cohort == 2) %>%
  Filter_Rows_Under_Max(100, FALSE) %>%
  arrange(ranking, desc(n_appointments), desc(n_pat)) %>%
  select(ranking, icd10, description, n_pat, n_appointments)
```

### Secondary Diagnoses
```{r Secondary Diag}

Secondary_Diags <- Overall_Cohort %>%
  inner_join(IP_Filtered, by = "hesid") %>%
  select(cohort, hesid, disdate, diag_02) %>%
  rename(value = diag_02) %>%
  filter(!is.na(value)) %>%
  mutate(icd10 = substring(value,1,4)) %>%
  inner_join(Full_Code_Lookup, by = "icd10") %>%
  select(cohort, hesid, icd10, description) %>%
  compute()

Secondary_Ranking <- Primary_Diags %>%
  count(hesid, cohort, icd10, description) %>%
  ungroup() %>%
  group_by(cohort, icd10, description) %>%
    summarise(n_pat = as.numeric(n()),
              n_appointments = as.numeric(sum(n))) %>%
  ungroup() %>%
  #Collect before running the section below otherwise it will rank things incorrectly
  collect() %>%
  #Rank ICD codes by number of appointments with them
  group_by(cohort) %>%
    arrange(desc(n_appointments), desc(n_pat)) %>%
    #If prior number of appointments and number of patients is the same set to 0, otherwise set to 1
    #Otherwise set to 1
    mutate(differ_prior = case_when(lag(n_appointments) == n_appointments & lag(n_pat) == n_pat ~ 0,
                                   TRUE ~ 1)) %>%
    #Cumsum the different prior to get rankings
    mutate(ranking = cumsum(differ_prior)) %>%
  ungroup() %>%
  select(-differ_prior)

Secondary_Diag_Adult_Rank <- Secondary_Ranking %>%
  filter(cohort == 1) %>%
  Filter_Rows_Under_Max(100, FALSE) %>%
  arrange(ranking, desc(n_appointments), desc(n_pat)) %>%
  select(ranking, icd10, description, n_pat, n_appointments) %>%
  collect()
  
Secondary_Diag_Non_Adult_Rank <- Secondary_Ranking %>%
  filter(cohort == 2) %>%
  Filter_Rows_Under_Max(100, FALSE) %>%
  arrange(ranking, desc(n_appointments), desc(n_pat)) %>%
  select(ranking, icd10, description, n_pat, n_appointments) %>%
  collect()

```

### Procedure Ranking
```{r Procedures}
Procedure_Names <- tbl(hes_con, in_schema(schema,"vantage_mapping_202111")) %>% 
  filter(str_detect(category, "proc")) %>%
  rename(opcs4 = code, desc4 = value) %>%
  select(opcs4, description = desc4) %>%
  compute

# Important to filter out Primary Opernosis as well as NAs and blanks
Procedures <- Overall_Cohort %>%
  inner_join(IP_Pivoted, by = "hesid") %>%
  filter(code_type == "oper") %>%
  select(cohort, hesid, disdate, code_type, code) %>%
  rename(value = code) %>%
  filter(!is.na(value)) %>% 
  filter(!value == "") %>%
  mutate(opcs4 = substring(value,1,4)) %>%
  left_join(Procedure_Names, by = "opcs4") %>%
  select(cohort, hesid, opcs4, description) %>%
  compute()

Procedure_Ranking <- Procedures %>% 
  count(hesid, opcs4, cohort, description) %>%
  ungroup() %>%
  group_by(cohort, opcs4, description) %>%
    summarise(n_pat = as.numeric(n()),
              n_appointments = as.numeric(sum(n))) %>%
  ungroup() %>%
  #Collect before running the section below otherwise it will rank things incorrectly
  collect() %>%
  #Rank ICD codes by number of appointments with them
  group_by(cohort) %>%
    arrange(desc(n_appointments), desc(n_pat)) %>%
    #If prior number of appointments and number of patients is the same set to 0, otherwise set to 1
    #Otherwise set to 1
    mutate(differ_prior = case_when(lag(n_appointments) == n_appointments & lag(n_pat) == n_pat ~ 0,
                                   TRUE ~ 1)) %>%
    #Cumsum the different prior to get rankings
    mutate(ranking = cumsum(differ_prior)) %>%
  ungroup() %>%
  select(-differ_prior)

Procedures_Adult_Rank <- Procedure_Ranking %>%
  filter(cohort == 1) %>%
  Filter_Rows_Under_Max(100, FALSE) %>%
  arrange(ranking, desc(n_appointments), desc(n_pat)) %>%
  select(ranking, opcs4, description, n_pat, n_appointments) %>%
  collect()
  
Procedures_Non_Adult_Rank <- Procedure_Ranking %>%
  filter(cohort == 2) %>%
  Filter_Rows_Under_Max(100, FALSE) %>%
  arrange(ranking, desc(n_appointments), desc(n_pat)) %>%
  select(ranking, opcs4, description, n_pat, n_appointments) %>%
  collect()

```


### Treatment Specialities
```{r}

Tretspef_Names <- tbl(hes_con, in_schema(schema,"vantage_mapping_202111")) %>% 
  filter(category == "tretspef") %>%
  rename(tretspef = code, description = value) %>%
  select(tretspef, description) %>%
  compute()

# Important to filter out Primary Diagnosis as well as NAs and blanks
Tretspefs <- Overall_Cohort %>%
  inner_join(IP_Filtered, by = "hesid") %>%
  select(cohort, hesid, disdate, tretspef) %>%
  inner_join(Tretspef_Names, by = "tretspef") %>%
  select(cohort, hesid, tretspef, description) %>%
  compute()

Treatspef_Ranking <- Tretspefs %>%
  count(hesid, cohort, description) %>%
  ungroup() %>%
  group_by(cohort, description) %>%
    summarise(n_pat = as.numeric(n()),
              n_appointments = as.numeric(sum(n))) %>%
  ungroup() %>%
  #Collect before running the section below otherwise it will rank things incorrectly
  collect() %>%
  #Rank ICD codes by number of appointments with them
  group_by(cohort) %>%
    arrange(desc(n_appointments), desc(n_pat)) %>%
    #If prior number of appointments and number of patients is the same set to 0, otherwise set to 1
    #Otherwise set to 1
    mutate(differ_prior = case_when(lag(n_appointments) == n_appointments & lag(n_pat) == n_pat ~ 0,
                                   TRUE ~ 1)) %>%
    #Cumsum the different prior to get rankings
    mutate(ranking = cumsum(differ_prior)) %>%
  ungroup() %>%
  select(-differ_prior)

Treatspef_Adult_Rank <- Treatspef_Ranking %>%
  filter(cohort == 1) %>%
  Filter_Rows_Under_Max(20, FALSE) %>%
  arrange(ranking, desc(n_appointments), desc(n_pat)) %>%
  select(ranking, description, n_pat, n_appointments) %>%
  collect()
  
Treatspef_Non_Adult_Rank <- Treatspef_Ranking %>%
  filter(cohort == 2) %>%
  Filter_Rows_Under_Max(20, FALSE) %>%
  arrange(ranking, desc(n_appointments), desc(n_pat)) %>%
  select(ranking, description, n_pat, n_appointments) %>%
  collect()

```

## Write each table as a tab

```{r}
write.xlsx(list("Primary Diagnosis Adult" = Primary_Diag_Adult_Rank, 
                "Primary Diagnosis Non-Adult" = Primary_Diag_Non_Adult_Rank, 
                "Secondary Diagnosis Adult" = Secondary_Diag_Adult_Rank, 
                "Secondary Diagnosis Non-Adult" = Secondary_Diag_Non_Adult_Rank, 
                "Procedures Adult" = Procedures_Adult_Rank, 
                "Procedures Non-Adult" = Procedures_Non_Adult_Rank, 
                "Tretspef Adult" = Treatspef_Adult_Rank, 
                "Tretspef Non-Adult" = Treatspef_Non_Adult_Rank), 
           "Output/Ranked_Analysis.xlsx")
```

